Copy number inference with InferCNV

### 2024-05-21 analyses/cell-type-ewings/template_notebooks/04-infercnv.Rmd
remove has_cnv umaps from template

reference_cells_only <- all_cnv_df |> 
  dplyr::filter(stringr::str_detect(subcluster, "reference")) |> 
  dplyr::mutate(has_cnv = NA,
                z_scaled_cnv = NA)

unknown_cells_only <- all_cnv_df |>
  dplyr::filter(stringr::str_detect(subcluster, "unknown")) |> 
  # get total number of chromosomes with cnvs per cell and then scale that number 
  dplyr::mutate(has_cnv = rowSums(across(has_cnv_cols)),
                z_scaled_cnv = scale(has_cnv))

# combine back together for plotting 
cnv_df <- bind_rows(reference_cells_only, unknown_cells_only)
```

@@

Below we will plot whether or not the gain or loss is present in each cell according to `InferCNV`.

```{r}
# umaps showing where cells are with gains/losses 
ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr8)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr8 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr12)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr12 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_dupli_chr1)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr1 gain") +
  theme_bw()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = has_loss_chr16)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr16 loss") +
  theme_bw()
```

We can also look at the proportion of the chromosome with the gain.

```{r}
ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr8)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr8 gain") +
  theme_bw() +
  scale_color_viridis_c()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr12)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr12 gain") +
  theme_bw() +
  scale_color_viridis_c()

ggplot(cnv_df, aes(x= UMAP1, y = UMAP2, color = proportion_scaled_dupli_chr1)) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(title = "Chr1 gain") +
  theme_bw() +
  scale_color_viridis_c()
```

>>>>

cnv_df <- all_cnv_df |> 
  dplyr::mutate(
    has_cnv = ifelse(
      stringr::str_detect(subcluster, "reference"), 
      NA_real_, 
      rowSums(across(has_cnv_cols))
    ),
    z_scaled_cnv = scale(has_cnv))
```

@@

First we will compare the proportions across each chromosome between cells we have labeled as tumor or normal. 
Then we will look specifically at the chromosomes commonly affected in Ewing sarcoma.

```{r}
# prepare a data frame with one row per chromosome per cell for plotting
proportion_df <- cnv_df |> 
  # select only relevant columns for plotting
  dplyr::select(barcodes, UMAP1, UMAP2, cnv_mean_classification, all_of(proportion_cols)) |> 
  tidyr::pivot_longer(cols = all_of(proportion_cols),
    names_to = "chromosome",
    values_to = "proportion") |> 
  # remove extra string in front of chr 
  dplyr::mutate(
    chromosome = stringr::str_remove(chromosome, "proportion_scaled_cnv_chr"),
    # turn into numeric for easy ordering in plot 
    chromosome = as.numeric(chromosome)
  )

```

```{r, fig.height=10, fig.width=10}
ggplot(proportion_df, aes(x = proportion, color = cnv_mean_classification)) +
  geom_density() +
  theme_bw() +
  facet_wrap(vars(chromosome), ncol = 3)
```

Below we plot only Chr1, Chr8, Chr12, and Chr16 and look at the proportion on a UMAP.

```{r}
ewings_chr_df <- proportion_df |> 
  dplyr::filter(chromosome %in% c(1,8,12,16))

ggplot(ewings_chr_df, aes(x = UMAP1, y=UMAP2 , color = proportion)) +
  geom_point(alpha = 0.5, size = 0.5) +
  theme_bw() +
  facet_wrap(vars(chromosome)) +
  scale_color_viridis_c()

```
